Modal Abstractions for Virtualizing Memory Addresses

### 2023-04-12 programlogic.tex
+

\begin{figure*}
  \begin{lstlisting}[language=Coq]
 Definition $\pfpointsto\locsf\vpts\qfrac\ppts$ :vProp $\Sigma$ := pmapsto_def_v (pfn w) (page_off w) q v.
 Definition $\vaddr\mapsto_{\textsf{t}}\{\textsf{q}\}\vpage$ :vProp $\Sigma$ := 
  $\exists_{\textsf{l4e l3e l2e l1}} \ldotp$  $\ulcorner$ aligned $\vaddr \urcorner \ast$ L4_L1_PointsTo(rtv maddr l4e l3e l2e l1e) $\ast$ addr_L1$(\vaddr,\textsf{l1e}) \mapsto_{p} \vpage$.
 Definition L4_L1_PointsTo (rtv maddr l4e l3e l2e l1e :word 64) : vProp $\Sigma$ :=
  (l4M52 maddr rtv) \$\sim$\ (l4off maddr rtv) $\mapsto_{a}$ {q1} l4e  $\ast$
  (l3M52 maddr l4e) \$\sim$\ (l3off maddr l4e)  $\mapsto_{a}$ {q2} l3e $\ast$ 
  (l2M52 maddr l3e) \$\sim$\ (l2off maddr l3e) $\mapsto_{a}$ {q3} l2e $\ast$
  (l1M52 maddr l2e) \$\sim$\ (l1off maddr l2e) $\mapsto_{a}$ {q4} l1e.
\end{lstlisting}
\caption{A Strong Virtual Points-to Relation}
\todo[inline,color=red]{Colin says: shouldn't ther be a fraction q in the body of the virtual pointsto def?
I know this is just a sketch that we then move away from, but it's worth getting accurate.}
\todo[inline,color=cyan]{Colin says: the virtual points-to here is not defined as a vprop! the body is an iProp,
containing an unbound variable rtv, which should be the cr3val taken as an argument. (Also, rtv doesn't make sense to me
as a shorthand, why not cr3val?). I'm also pretty sure the maddr in the virtual points-to definition should be va.}
  \label{fig:strongvirtualpointsto}
\end{figure*}

@@

\todo[inline,color=red]{Colin says: Why are we talking about assertions for address translations here? It seems like we should
\todo[inelin]{Colin says: There's no explanation of the split 52+12 \emph{notation} here. If we added such an explanation
(maybe to where physical points-tos are introduced?) and moved this discussion to \ref{sec:overly-restrictive}, this is otherwise
a nice explanation}
discuss physical memory pointsto assertions in general first, then talk about this in \ref{sec:overly-restrictive}}
Given the definition of physical page-pointsto assertion and the root address of virtual-address space as shown in Figure \ref{fig:pagetable}, one can build the physical address-translation for a virtual address (e.g. \textsf{va}) via abstracting the L4-L1 table traversal as the following:
\begin{itemize}
  \item Level-4 Translation (L4): Performs the first level address translation to get the L4 entry (L4E) in Figure \ref{fig:pagetables} by using the masks l4M52 and l4off with \textsf{rtv} virtual base address to get the starting address of the L4 table (L4T), and locate the entry entry amongst 512 (q1) respectively.
    \begin{lstlisting}[language=Coq]
      $\hbox{(\TirNameStyle{{L4translate}})} \quad$ (l4M52 maddr rtv) \$\sim$\ (l4off maddr rtv) $\mapsto_{a}$ {q1} l4e 
    \end{lstlisting}
 \item Level-3 Translation (L3): Performs the second level address translation to get the L3 entry (L3E) in Figure \ref{fig:pagetables} by using the masks l3M52 and l3off with L4 entry (l4e) to get the starting address of the L3 table (L3T), and locate the entry amongst 512 (q2) respectively.
    \begin{lstlisting}[language=Coq]
    $\hbox{(\TirNameStyle{{L3translate}})}$ (l3M52 maddr l4e) \$\sim$\ (l3off maddr l4e)  $\mapsto_{a}$ {q2} l3e
    \end{lstlisting}
  \item Level-2 Translation (L2): Performs the second level address translation to get the L2 entry (L2E) in Figure \ref{fig:pagetables} by using the masks l2M52 and l2off with L3 entry (l3e) to get the starting address of the L2 table (L2T), and locate the entry entry amongst 512 (q3) respectively.
\begin{lstlisting}[language=Coq]
    $\hbox{(\TirNameStyle{{L2translate}})}$ (l2M52 maddr l3e) \$\sim$\ (l2off maddr l3e)  $\mapsto_{a}$ {q3} l2e
    \end{lstlisting}
  \item Level-1 Translation (L1): Performs the second level address translation to get the L1 entry (L1E) in Figure \ref{fig:pagetables} by using the masks l1M52 and l1off with L2 entry (l2e) to get the starting address of the L1 table (L1T), and locate the entry entry amongst 512 (q4) respectively.
   \begin{lstlisting}[language=Coq]
    $\hbox{(\TirNameStyle{{L1translate}})}$ (l1M52 maddr l2e) \$\sim$\ (l1off maddr l2e)  $\mapsto_{a}$ {q4} l1e
    \end{lstlisting}
  \item Page Address Level Translation: Final computed physical page-address ($\textsf{addr\_L1}(\vaddr$,$\textsf{l1e}$)) points-to the value stored in the address ($\vpage$).
   \begin{lstlisting}[language=Coq]
    $\hbox{(\TirNameStyle{{PageLabelAccess}})} \qquad$ addr_L1($\vaddr,\textsf{l1e}$) $\mapsto_{p} \vpage$.
    \end{lstlisting} 
\end{itemize}

>>>>

\begin{figure*}
  \begin{lstlisting}[language=Coq]
 Definition $\pfpointsto\locsf\vpts\qfrac\ppts$ :vProp $\Sigma$ := pmapsto_def_v (pfn w) (page_off w) q v.
 Definition $\vaddr\mapsto_{\textsf{t}}\{\textsf{q}\}\; \vpage$ : vProp $\Sigma$ := 
  $\exists_{\textsf{l4e l3e l2e l1}} \ldotp$  $\ulcorner$ aligned $\vaddr \urcorner \ast$ L4_L1_PointsTo(rtv maddr l4e l3e l2e l1e) $\ast$ addr_L1$(\vaddr,\textsf{l1e}) \mapsto_{p} \vpage$.
 Definition L4_L1_PointsTo (rtv maddr l4e l3e l2e l1e :word 64) : vProp $\Sigma$ :=
  (l4M52 maddr rtv) \$\sim$\ (l4off maddr rtv) $\mapsto_{a}$ {q1}  l4e  $\ast$
  (l3M52 maddr l4e) \$\sim$\ (l3off maddr l4e)  $\mapsto_{a}$ {q2}  l3e $\ast$ 
  (l2M52 maddr l3e) \$\sim$\ (l2off maddr l3e) $\mapsto_{a}$ {q3}  l2e $\ast$
  (l1M52 maddr l2e) \$\sim$\ (l1off maddr l2e) $\mapsto_{a}$ {q4}  l1e.
\end{lstlisting}
\caption{A Strong Virtual Points-to Relation}
\todo[inline,color=red]{Colin says: shouldn't ther be a fraction q in the body of the virtual pointsto def?
I know this is just a sketch that we then move away from, but it's worth getting accurate.}
\todo[inline,color=cyan]{Colin says: the virtual points-to here is not defined as a vprop! the body is an iProp,
containing an unbound variable rtv, which should be the cr3val taken as an argument. (Also, rtv doesn't make sense to me
as a shorthand, why not cr3val?). I'm also pretty sure the maddr in the virtual points-to definition should be va.}
  \label{fig:strongvirtualpointsto}
\end{figure*}

@@

\todo[inline,color=red]{Colin says: Why are we talking about assertions for address translations here? It seems like we should
\todo[inelin]{Colin says: There's no explanation of the split 52+12 \emph{notation} here. If we added such an explanation
(maybe to where physical points-tos are introduced?) and moved this discussion to \ref{sec:overly-restrictive}, this is otherwise
a nice explanation}
discuss physical memory pointsto assertions in general first, then talk about this in \ref{sec:overly-restrictive}}
Given the definition of physical page-pointsto assertion and the root address of virtual-address space as shown in Figure \ref{fig:pagetable}, one can build the physical address-translation for a virtual address (e.g. \textsf{va}) via abstracting the L4-L1 table traversal as the following:
\begin{itemize}
  \item Level-4 Translation (L4): Performs the first level address translation to get the L4 entry (L4E) in Figure \ref{fig:pagetables} by using the masks l4M52 and l4off with \textsf{rtv} virtual base address to get the starting address of the L4 table (L4T), and locate the entry entry amongst 512 (q1) respectively.
    \begin{lstlisting}[language=Coq]
      $\hbox{(\TirNameStyle{{L4translate}})} \quad$ (l4M52 maddr rtv) \$\sim$\ (l4off maddr rtv) $\mapsto_{a}$ {q1} l4e 
    \end{lstlisting}
 \item Level-3 Translation (L3): Performs the second level address translation to get the L3 entry (L3E) in Figure \ref{fig:pagetables} by using the masks l3M52 and l3off with L4 entry (l4e) to get the starting address of the L3 table (L3T), and locate the entry amongst 512 (q2) respectively.
    \begin{lstlisting}[language=Coq]
    $\hbox{(\TirNameStyle{{L3translate}})}$ (l3M52 maddr l4e) \$\sim$\ (l3off maddr l4e)  $\mapsto_{a}$ {q2} l3e
    \end{lstlisting}
  \item Level-2 Translation (L2): Performs the second level address translation to get the L2 entry (L2E) in Figure \ref{fig:pagetables} by using the masks l2M52 and l2off with L3 entry (l3e) to get the starting address of the L2 table (L2T), and locate the entry entry amongst 512 (q3) respectively.
\begin{lstlisting}[language=Coq]
    $\hbox{(\TirNameStyle{{L2translate}})}$ (l2M52 maddr l3e) \$\sim$\ (l2off maddr l3e)  $\mapsto_{a}$ {q3} l2e
    \end{lstlisting}
  \item Level-1 Translation (L1): Performs the second level address translation to get the L1 entry (L1E) in Figure \ref{fig:pagetables} by using the masks l1M52 and l1off with L2 entry (l2e) to get the starting address of the L1 table (L1T), and locate the entry entry amongst 512 (q4) respectively.
   \begin{lstlisting}[language=Coq]
    $\hbox{(\TirNameStyle{{L1translate}})}$ (l1M52 maddr l2e) \$\sim$\ (l1off maddr l2e)  $\mapsto_{a}$ {q4} l1e
    \end{lstlisting}
  \item Page Address Level Translation: Final computed physical page-address ($\textsf{addr\_L1}(\vaddr$,$\textsf{l1e}$)) points-to the value stored in the address ($\vpage$).
   \begin{lstlisting}[language=Coq]
    $\hbox{(\TirNameStyle{{PageLevelAccess}})} \qquad$ addr_L1($\vaddr,\textsf{l1e}$) $\mapsto_{p} \vpage$.
    \end{lstlisting} 
\end{itemize}
